On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-01-15 manuscript.Rmd
more edits; version sent to co-authors 15 Jan

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from cuvette data
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to drought
\t4. Acclimation and instantaneous response to temperature
\t5. Other factors
VIII. Conclusions", header=F)

@@

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research. Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers.

@@

\ 
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.  Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^ was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.

@@

**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but some data occurs outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cut,isol~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

# III. Problems with estimating minimum conductance from cuvette data

@@

\
It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves. Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).

@@

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry). In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq. 1 is lower (Fig. 3c).

@@

**Figure 3**. Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases. (a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.  Vertical lines are 95% confidence intervals. The grey line is a fitted loess smoother with 95% confidence interval. Note the wide confidence intervals, and frequent negative values. (b) Similar to (a), but using the published compilation by @miner2017. The grey line is a fitted loess smoother with 95% confidence interval. (c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e. right-hand side of the equation being fitted) is lower.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake, @urban2017].

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
When plants are sufficiently water-stressed that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 5).

@@

**Figure 5**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation. (c) Progression of percent loss conductivity (PLC) of the xylem. If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer (or nearly 40%).

@@

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation explains only some of the wide range of g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition where no photosynthesis takes place. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to any other measure of minimum conductance. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.

@@

A few studies have directly quantified the stomatal component of g~min~. A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.

@@

\
While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species *wow rewrite*. Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude. This discussion is directly relevant to the use of g~min~ in models, because if intra-specific variation is large, it complicates model parameterization.

>>>>

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from regression
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to the environment
VIII. Conclusions", header=F)

@@

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge. Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers.

@@

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.  Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area. Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.

@@

\
(Figure 1 here)

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cut~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

# III. Problems with estimating minimum conductance from regression

@@

\
It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated. Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves. Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the g~1~ parameter was unaffected).

@@

\
(Figure 2 here)

\
**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry). In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because there was low variation in environmental conditions (humidity, light, temperature). We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq. 1 is lower (Fig. 3c).

@@

\
(Figure 3 here)

\
**Figure 3**. Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases. (a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.  Vertical lines are 95% confidence intervals. The grey line is a fitted loess smoother with 95% confidence interval. Note the wide confidence intervals, and frequent negative values. (b) Similar to (a), but using the published compilation by @miner2017. The grey line is a fitted loess smoother with 95% confidence interval. (c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e. right-hand side of the equation being fitted) is lower.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favourable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with the photon flux density (PPFD), as is typical in leaf gas exchange data. When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PPFD.

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

\
(Figure 4 here)

\
**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
When plants are sufficiently water-stressed so that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 5).

@@

\
(Figure 5 here)

\
**Figure 5**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation. (c) Progression of percent loss conductivity (PLC) of the xylem. If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer.

@@

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cut~ measured across different plant species. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to any other measure of minimum conductance. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~, and should be modelled in a separate framework that is yet to be identified.

@@

In order to understand variation in g~min~, we have to know more about the role of the cuticle vs. leaky or incompletely closed stomata. A few studies have directly quantified the stomatal component of g~min~. A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.

@@

\
While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species. Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude. This discussion is directly relevant to the use of g~min~ in models, because if intra-specific variation is large, it complicates model parameterization.
